---
title: "Conditional controls and screen output"
author: "Julien Barnier"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Conditional controls and screen output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Conditional controls

Sometimes you don't want your user to be able to go to the next screen if a condition based on inputs is not met. This can be done in `shinyglide` with the `next_condition` argument of the `screen` function. This option takes exactly the same syntax has the one used in `shiny::conditionalPanel`.

For example, in the following app, the *Next* button is disabled until the *n* value is greater then zero :

```{r eval = FALSE}
ui <- fluidPage(
  glide(
    screen(
      next_condition = "input.n > 0",
      
      p("Please choose a value for n :"),
      numericInput("n", "n :", value = 0)
    ),
    screen(
      p("Here is your plot :"),
      plotOutput("plot")
    )
  )
)

server <- function(input, output, session) {
  output$plot <- renderPlot({
    hist(rnorm(input$n), main = paste("n =", input$n))
  })
}

shinyApp(ui, server)
```

You can do the same for the *Back* button with the `previous_condition` argument.

By default the button is disabled when the condition is not met (more precisely, the button gets a `disabled = "disabled"` argument and a `disabled` CSS class). If you prefer to hide it entirely, you can add `disable_type = "hide"` to your `glide()` definition :

```{r eval = FALSE}
ui <- fluidPage(
  glide(
    disable_type = "hide",
    
    screen(
      next_condition = "input.n > 0",
      
      p("Please choose a value for n :"),
      numericInput("n", "n :", value = 0)
    ),
    screen(
      p("Here is your plot :"),
      plotOutput("plot")
    )
  )
)

server <- function(input, output, session) {
  output$plot <- renderPlot({
    hist(rnorm(input$n), main = paste("n =", input$n))
  })
}

shinyApp(ui, server)
```

## Screen output

Sometimes you want to show a screen only if certain conditions are met based on inputs. For example, if a user chooses a value which is not present in a dataset, you would like to display a specific screen to alert him.

This is possible by using a `screenOutput` call in you app UI, associated to a `renderUI` in your app server. If `renderUI` returns `NULL`, the screen will be hidden.

Here is an example :

```{r eval = FALSE}
ui <- fluidPage(
  glide(
    disable_type = "hide",
    
    screen(
      p("Do you want to see the next screen ?"),
      checkboxInput("check", "Yes, of course !", value = FALSE)
    ),
    screenOutput("check_screen"),
    screen(
      p("And this is the last screen")
    )
  )
)

server <- function(input, output, session) {
  output$check_screen <- renderUI({
    if(!input$check) return(NULL)
    p("Here it is !")
  })
  outputOptions(output, "check_screen", suspendWhenHidden = FALSE)
}

shinyApp(ui, server)
```

**Important :** for the `screenOutput` to work, it is mandatory to add the following line in your app server :

```{r eval=FALSE}
outputOptions(output, "your_screen_id", suspendWhenHidden = FALSE)
```

Otherwise, it won't be updated when hidden.

Sometimes the `screenOutput` computation takes some time. In these cases, with the default controls, the *Next* button is disabled and a *Loading* message is shown until the next screens computations is done. You can see it in the following example, which is the same as the previous one but with an articial 2 seconds waiting time :

```{r eval = FALSE}
ui <- fluidPage(
  glide(
    disable_type = "hide",
    
    screen(
      p("Do you want to see the next screen ?"),
      checkboxInput("check", "Yes, of course !", value = FALSE)
    ),
    screenOutput("check_screen"),
    screen(
      p("And this is the last screen")
    )
  )
)

server <- function(input, output, session) {
  output$check_screen <- renderUI({
    Sys.sleep(2)
    if(!input$check) return(NULL)
    p("Here it is !")
  })
  outputOptions(output, "check_screen", suspendWhenHidden = FALSE)
}

shinyApp(ui, server)
```


### Customizing loading

You can customize the loading label and animation with options passed to `glide()` :

- `loading_class` (default value : `"loading"`) : a CSS class passed to the *Next* button when in loading state
- `loading_label` : the label of the *Next* button while in loading state. This can be any HTML, and you can insert a `span(class = "shinyglide-spinner")` in it if you want to display the defautl animated spinner somewhere.

You can take a look at the [custom controls sample app](https://data.nozav.org/app/shinyglide/04_custom_controls/) for an example of loading customization with [custom controls](https://juba.github.io/shinyglide/articles/custom_controls.html) and CSS.


